---
title: "p8105_hw2_xw2961"
output: github_document
---

# Load package

```{r}
library(tidyverse)
library(readxl)
library(tidyr)
library(dplyr)
```

# Problem 1

```{r}
NYC_subway = read.csv("~/Desktop/Columbia Year2/Data Science/HW/HW2/p8105_hw2/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
```

# Problem 2

-   import and clean the dataset (Mr. Trash Wheel)

```{r}
Mr_trash_wheel = read_excel("~/Desktop/Columbia Year2/Data Science/HW/HW2/p8105_hw2/202409 Trash Wheel Collection Data.xlsx",
range="A2:N655")%>%
  janitor:: clean_names()%>% 
  mutate(sports_balls = round(as.numeric(sports_balls),0))%>% 
  filter(!is.na(dumpster)) %>% mutate(year=as.character(year))
```

-   import and clean the dataset (Professor Trash Wheel)

```{r}
Prof_trash_wheel = read_excel("~/Desktop/Columbia Year2/Data Science/HW/HW2/p8105_hw2/202409 Trash Wheel Collection Data.xlsx", 
sheet = "Professor Trash Wheel", range="A2:M123") %>%
  janitor::clean_names() %>%
  filter(!is.na(dumpster)) %>%
  filter(!is.na(date)) %>% mutate(year=as.character(year))
```

-   import and clean the dataset (Professor Gwynnda Trash Wheel)

```{r}
gwynnda_trash_wheel = read_excel("~/Desktop/Columbia Year2/Data Science/HW/HW2/p8105_hw2/202409 Trash Wheel Collection Data.xlsx",
sheet = "Gwynnda Trash Wheel", range="A2:L266")%>%
  janitor::clean_names() %>%
    filter(!is.na(dumpster)) %>%
  filter(!is.na(date)) %>% mutate(year=as.character(year))
```

-   merge the datasets

```{r}
  combined_trash = bind_rows(Mr_trash_wheel,Prof_trash_wheel,gwynnda_trash_wheel) 
```

* This combined dataset contains dumpster, Year, Month, Date, Weight, Volume, Plastic Bottles, Polystyrene, Cigarette Butts, Glass Bottles, Plastic Bags, Wrappers, Sports Balls, Homes Powered. The combined dataset contains 1,032 observations.

-   Calculate the total weight of trash collected by Professor Trash Wheel is 
`r sum(Prof_trash_wheel$weight_tons, na.rm=TRUE)`


-   Calculate the total number of cigarette butts collected by Gwynnda in June of 2022 is 
`r gwynnda_trash_wheel %>% filter(month=="June" & year==2022) %>% summarise(total = sum(cigarette_butts, rm.na=TRUE)) %>% pull(total)`


# Problem 3
* import, clean, tidy, and otherwise wrangle each of these datasets
```{r}
bakers = read.csv("gbb_datasets/bakers.csv") %>% 
  janitor::clean_names() %>% 
  mutate(baker_occupation=na_if(baker_occupation,""), hometown=na_if(hometown,"") ) %>% drop_na() #turn missing into N/A and drop the cells with N/A
  
bakes   = read.csv("gbb_datasets/bakes.csv")%>% 
  janitor::clean_names()%>%
  mutate(signature_bake=na_if(signature_bake,""), show_stopper=na_if(show_stopper,"")) %>% drop_na()

results = read.csv("gbb_datasets/results.csv", skip=2)%>% 
  janitor::clean_names()%>%
  mutate(result=na_if(result,"gbb_datasets/results.csv")) %>% drop_na()
```

* Anti joint to check for completeness and correctness across datasets
```{r}
# Keep Firstname only for the bakers dataset
bakers  = bakers %>% mutate(baker=sub(" .*", "", baker_name)) %>% select(-baker_name) %>% relocate(baker)
results = results %>% relocate(baker)
bakes   = bakes %>% relocate(baker)

# Check bakers that are in bakers but not in bakes
unmatched_bakers_in_bakes <- anti_join(bakers, bakes, by = "baker")
unmatched_bakers_in_results <- anti_join(bakers, results, by = "baker")

```

* participants `r unmatched_bakers_in_bakes$baker` that appeared in bakes dataset do not have a matching record in bakers dataset
* participants `r unmatched_bakers_in_results$baker` that appeared in bakes dataset do not have a matching record in bakers dataset

* Merge dataset & put in meaningful order

```{r}
final_dataset =
  bakers %>% 
  left_join(results, by = c("baker", "series")) %>% 
  left_join(bakes, by = c("baker", "series", "episode")) %>% 
  select(series, episode, baker, baker_age, baker_occupation, hometown,
         signature_bake, show_stopper, technical, result) %>% 
  arrange(series, episode, baker)
write_csv(final_dataset, "gbb_datasets/final_dataset.csv")
```

* Data cleaning process

First we import the datasets using `read.csv` and utilize a consistent column naming convention with `janitor::clean_names()`. 
